Temporal Evolution of Erosion in Pore-Networks:\\ From Homogenization to Instability

### 2022-03-06 response_submission/supp.tex
Updates from Overleaf

\usepackage[usenames, dvipsnames]{color}
% Color Edits
\newcommand{\EM}[1]{\noindent \color{red} (AA: #1)\normalcolor}
\newcommand{\AZ}[1]{\noindent \color{blue} (AZ: #1)\normalcolor}

\usepackage{braket}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}
\usepackage[usenames, dvipsnames]{color}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

\renewcommand{\thepage}{S\arabic{page}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\theequation}{S\arabic{equation}}
\def\lp{\left(}
\def\rp{\right)}
\def\lb{\left[}
\def\rb{\right]}
\begin{document}

@@

\title{Temporal Evolution of Flow in Pore-Networks:\\ From Homogenization to Instability}% Force line breaks with \\
% \thanks{A footnote to the article title}%

\author{Ahmad Zareei}
\affiliation{ School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, 02148}%\altaffiliation[]{SEAS, Pierce Hall}%Lines break automatically or can be forced with \\
\author{Pan Deng}%
\affiliation{ School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, 02148}%\altaffiliation[]{SEAS, Pierce Hall}%Lines break automatically or can be forced with
\author{Ariel Amir}%
\email{arielamir@seas.harvard.edu}
\affiliation{ School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, 02148}%\altaffiliation[]{SEAS, Pierce Hall}%Lines break automatically or can be forced with
% need to figure out how to reference dagger

@@

This condition guarantees that at each step a small amount of material is eroded and there is no sudden change in the network. We further test the convergence by decreasing $\max(\Delta r_{ij})$ to half and we observe that the average relative change in the flux vector is $\approx 1.2\%$, and the PDFs remain intact without any notable change. The network size for the 2d/3d networks is $N_x=100,N_y=50$/$N_x=50,N_y=12, N_z=12$ unless mentioned otherwise. In either case, the network has an order of $10^3$ edges. The code is publicly available in a GitHub repository \cite{githubrepo}.

%
% \begin{figure}[h]
%     % \centering
%     \includegraphics[width = 0.75\textwidth]{Figs/Fig2_random-low.png}
%     \caption{Ahmad's version: Erosion in a topologically random network of pipes. The first row shows the initial condition at $t=0$. Each row afterward corresponds to the simulation result after $N$ steps such that $\langle r_{t=N}\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion. The first column is a snapshot of the pore network, the second column is the PDF of normalized fluid flux $q/\langle q \rangle$, and the last column is the PDF of normalized radius $r/\langle r\rangle$.}\label{fig:fig2}
% \end{figure}
%

@@

\newpage
\section{Robustness to topology and initial condition}
% 
In order to check the robustness of our result with respect to the topology of the network, we run our simulations on a three-dimensional random network with Voronoi cell initialization of nodes in space (Fig. \ref{SIfig:fig2-3d}), and also for a two dimensional topologically ordered network with a diamond grid (Fig. \ref{fig:fig2-diamond}), where  both networks are initialized with a uniform random distribution for the diameter of the tubes. We find that regardless of the topology of the network, an erosion dynamics with $n>3$ results in a homogenized network, while $n<3$ results in the channelization instability. We further check the effect of initial randomness on the fate of the network. We use a two-dimensional diamond-grid network with a narrow uniform initial distribution of tube diameters around an average diameter $d_0$ with only a very small variation ($3\%$), i.e., the tube diameters are sampled from $\mathcal{U}(d_0(1-\epsilon), d_0(1+\epsilon))$. We again find that for the networks with $n<3$ channels are formed, while for $n>3$ the network stays homogenized.  
%
\begin{figure}[htp]
     % \centering
      \includegraphics[width = 1.0\textwidth]{FigS2.pdf}
     \caption{Erosion in a topologically random 3D network of tubes with $N_x=20, N_y=12, N_z=12$ points for Voronoi cell initialization and an initial uniform broad distribution of tube diameters randomly sampled from $\mathcal{U}(1,14)$. Snapshots of the network, PDF of normalized fluid flux $q/\langle q \rangle$, and normalized edge radius distribution $r/\langle r \rangle$ at the initial time $t=0$, and also after $N$ erosion steps for different powers of erosion $n$ are shown. We stop the erosion after $N$ steps such that $\langle r\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion.}\label{SIfig:fig2-3d}
 \end{figure}
%
\begin{figure}[htp]
    % \centering
 \includegraphics[width=0.75\textwidth]{FigS3.pdf} % Fig2-Sdiamond.pdf
     \caption{Erosion in a diamond grid network with $N_x=100, N_y=50$ randomly distributed nodes and an initially broad  distribution of tube diameter randomly sampled from $\mathcal{U}(1,14)$. The initial condition is shown with the label $t=0$ in the first row. Each row afterward corresponds to the simulation result after $N$ steps such that $\langle r\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$ or twice the initial average radius. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion. The first column is a snapshot of the pore network, the second column is the PDF of normalized fluid flux $q/\langle q \rangle$, and the last column is the PDF of normalized radius $r/\langle r\rangle$.}\label{fig:fig2-diamond}% 
\end{figure}
%
\begin{figure}[htp]
    % \centering
 \includegraphics[width=0.72\textwidth]{FigS4.pdf}
     \caption{Erosion in a structured diamond grid network with $N_x=50, N_y=50$ and an initially narrow distribution of tube diameters randomly sampled from  $\mathcal{U}\left(d_0(1-\epsilon), d_u(1+\epsilon)\right) $ where  $d_0$ is the average diameter and $\epsilon=0.03$. The initial condition is shown with the label $t=0$ in the first row. Each row afterward corresponds to the simulation result after $N$ steps such that $\langle r_{t=N}\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$ or twice the initial average radius. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion. The first column is a snapshot of the pore network, the second column is the PDF of normalized fluid flux $q/\langle q \rangle$, and the last column is the PDF of normalized radius $r/\langle r\rangle$.}\label{fig:fig2}% 
\end{figure}

@@

% \clearpage
\bibliography{ref}% Produces the bibliography via BibTeX.

>>>>

\usepackage[usenames, dvipsnames]{color}
% Color Edits

\usepackage{braket}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}
\usepackage[usenames, dvipsnames]{color}
\usepackage{subfig}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

\usepackage{xcolor}
\newcommand{\DP}[1]{\noindent \color{cyan} (DP: #1)\normalcolor}
% \newcommand{\AZ}[1]{\noindent \color{purple} (AZ: #1)\normalcolor}
\newcommand{\add}[1]{\noindent \color{blue} #1 \normalcolor}
\newcommand{\del}[1]{\noindent \color{red} \st{#1}\normalcolor}
\newcommand{\AZ}[1]{\noindent \color{magenta} (AZ: #1)\normalcolor}
\renewcommand{\thepage}{S\arabic{page}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\theequation}{S\arabic{equation}}
\def\lp{\left(}
\def\rp{\right)}
\def\lb{\left[}
\def\rb{\right]}
\begin{document}

@@

\title{Temporal Evolution of Erosion in Pore-Networks:\\ From Homogenization to Instability}% Force line breaks with \\
% \thanks{A footnote to the article title}%

\author{Ahmad Zareei}
\affiliation{ School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, 02148}%\altaffiliation[]{SEAS, Pierce Hall}%Lines break automatically or can be forced with \\
\author{Deng Pan}%
\affiliation{ School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, 02148}%\altaffiliation[]{SEAS, Pierce Hall}%Lines break automatically or can be forced with
\author{Ariel Amir}%
\email{arielamir@seas.harvard.edu}
\affiliation{ School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, 02148}%\altaffiliation[]{SEAS, Pierce Hall}%Lines break automatically or can be forced with
% need to figure out how to reference dagger

@@

This condition guarantees that at each step a small amount of material is eroded and there is no sudden change in the network. We further test the convergence by decreasing $\max(\Delta r_{ij})$ to half and we observe that the average relative change in the flux vector is $\approx 1.2\%$, and the PDFs remain intact without any notable change. The network size for the 2d/3d networks is $N_x=100,N_y=50$/$N_x=50,N_y=12, N_z=12$ unless mentioned otherwise. In either case, the network has an order of $10^3$ edges. \add{Note that the physical assumption of low Reynolds regime Re$\ll$1 at all times ensures that the Poisoulle flow assumption remains valid. Additional dimensionless transport numbers such as the P\'eclet number (ratio of advection to diffusion) or the Damkohler number (ratio of chemical reaction rate to the transport phenomena rate) are assumed to be respectively very large and very small, resulting in advection to be much stronger than diffusion and chemical reaction.  It is to be noted that considering diffusion or chemical reactions might change the dynamics of erosion as it requires tracking a secondary concentration field that might affect the local erosion law.} The code is publicly available in a GitHub repository \cite{githubrepo}.
% \begin{figure}[h]
%     % \centering
%     \includegraphics[width = 0.75\textwidth]{Figs/Fig2_random-low.png}
%     \caption{Ahmad's version: Erosion in a topologically random network of pipes. The first row shows the initial condition at $t=0$. Each row afterward corresponds to the simulation result after $N$ steps such that $\langle r_{t=N}\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion. The first column is a snapshot of the pore network, the second column is the PDF of normalized fluid flux $q/\langle q \rangle$, and the last column is the PDF of normalized radius $r/\langle r\rangle$.}\label{fig:fig2}
% \end{figure}
\newpage

@@

\newpage
\section{Robustness to topology, initial condition, and boundary condition}
In order to check the robustness of our results with respect to the topology of the network, we run our simulations on a three-dimensional random network with Voronoi cell initialization of nodes in space (Fig. \ref{SIfig:fig2-3d}), and also for a two dimensional topologically ordered network with a diamond grid (Fig. \ref{fig:fig2-diamond}), where  both networks are initialized with a uniform random distribution for the diameter of the tubes. We find that regardless of the topology of the network, an erosion dynamics with $n>3$ results in a homogenized network, while $n<3$ results in the channelization instability. We further check the effect of initial randomness on the fate of the network. We use a two-dimensional diamond-grid network with a narrow uniform initial distribution of tube diameters around an average diameter $d_0$ with only a very small variation ($3\%$), i.e., the tube diameters are sampled from a uniform distribution with $\mathcal{U}(d_0(1-\epsilon), d_0(1+\epsilon))$. We again find that for the networks with $n<3$ channels are formed, while for $n>3$ the network stays homogenized. \add{In addition to the network topology and initial condition, we test the robustness of our results to changes in the boundary condition. We choose a circular domain, with the central node being the inlet and all the nodes at the outer boundary being outlets. Similar to the transition reported for the square
domain, we find a phase transition at the predicted power which shows that our analysis is independent of the geometry (Fig.~\ref{SIfig:fig2-3d_circular}). }
\begin{figure}[htp]
     % \centering
      \includegraphics[width = 1.0\textwidth]{FigS2.pdf}
     \caption{Erosion in a topologically random 3D network of tubes with $N_x=20, N_y=12, N_z=12$ points for Voronoi cell initialization and an initial uniform broad distribution of tube diameters randomly sampled from a uniform distribution with $\mathcal{U}(1,14)$. Snapshots of the network, PDF of normalized fluid flux $q/\langle q \rangle$, and normalized edge radius distribution $r/\langle r \rangle$ at the initial time $t=0$, and also after $N$ erosion steps for different powers of erosion $n$ are shown. We stop the erosion after $N$ steps such that $\langle r\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$. The erosion law is based on $dr/dt = \alpha q/r^n$ where different powers of $n$ correspond to different models of erosion.}\label{SIfig:fig2-3d}
\begin{figure}[htp]
    % \centering
 \includegraphics[width=0.75\textwidth]{FigS3.pdf} % Fig2-Sdiamond.pdf
     \caption{Erosion in a diamond grid network with $N_x=100, N_y=50$ randomly distributed nodes and an initially broad  distribution of tube diameter randomly sampled from a uniform distribution with $\mathcal{U}(1,14)$. The initial condition is shown with the label $t=0$ in the first row. Each row afterward corresponds to the simulation result after $N$ steps such that $\langle r\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$ or twice the initial average radius. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion. The first column is a snapshot of the pore network, the second column is the PDF of normalized fluid flux $q/\langle q \rangle$, and the last column is the PDF of normalized radius $r/\langle r\rangle$.}\label{fig:fig2-diamond}% 
\begin{figure}[htp]
    % \centering
 \includegraphics[width=0.72\textwidth]{FigS4.pdf}
     \caption{Erosion in a structured diamond grid network with $N_x=50, N_y=50$ and an initially narrow distribution of tube diameters randomly sampled from a uniform distribution with  $\mathcal{U}\left(d_0(1-\epsilon), d_u(1+\epsilon)\right) $ where  $d_0$ is the average diameter and $\epsilon=0.03$. The initial condition is shown with the label $t=0$ in the first row. Each row afterward corresponds to the simulation result after $N$ steps such that $\langle r_{t=N}\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$ or twice the initial average radius. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion. The first column is a snapshot of the pore network, the second column is the PDF of normalized fluid flux $q/\langle q \rangle$, and the last column is the PDF of normalized radius $r/\langle r\rangle$.}\label{fig:fig2}% 
\begin{figure}[htp]
     % \centering
      \includegraphics[width = 0.70\textwidth]{./circular3.pdf}
     \caption{\textcolor{blue}{Erosion in a topologically random 3D network of tubes with a circular topology with $N = 3000 $ randomly distributed nodes and an initial uniform broad distribution of tube diameters randomly sampled from a uniform distribution with $\mathcal{U}(1,14)$. Shown are snapshots of the network, PDF of normalized fluid flux $q/\langle q \rangle$, and normalized edge radius distribution $r/\langle r \rangle$ at the initial time $t=0$, and also after $N$ erosion steps for different powers of erosion $n$. We stop the erosion after $N$ steps such that $\langle r\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion.}}\label{SIfig:fig2-3d_circular}

\clearpage
\section{Erosion with a threshold}
%
\add{In order to make our erosion model more realistic, we modify the model as  
%
\begin{align}
    \frac{dr}{dt } = \begin{cases} q^m/r^n - \beta & q^m/r^n\geq \beta\\
    0 & q^m/r^n<\beta
    \end{cases} \label{eq:thresh}
\end{align}
%
where $\beta$ is the threshold for the onset of erosion. We pick $\beta = 0.25\langle q^m/r^n\rangle $ and $\beta = 0.9 \langle q^m/r^n\rangle$ corresponding to a small and a large threshold value. In both cases, we ran our simulations for different values of $m$ and $n$, and found that different behaviors of homogenization and channelization persists (see Fig. \ref{fig:figthresh1}). The result can be explained by applying the general condition for the phase transition as discussed in the main text. Following the same argument, it can be shown that the homogenization condition (Eq. 3 in the main text) is equivalent to the condition 
%
\begin{align}
4m-n-1 < -\frac{\beta}{q^m/r^n}.    
\end{align}
%
which can be found using $g(r) = ({\pi \Delta p}/{8\mu L})^{m} r^{m-n} - \beta $ assuming a constant pressure boundary condition. Note that the condition above depends on the value of $q^m/r^n$ which varies over the network edges. Taking the mean value of $q^m/r^n$ to find an average value for the transition condition, the transition line becomes $4m-n=0.75$ and $4m-n=0.1$ for the small and large thresholding value respectively. As a result the thresholding condition slightly shifts the intercept of the transition line while the slope remains the same. In Fig.~\ref{fig:figthresh1}, we plot the new transition lines, where the prediction for the phase boundaries matches well with the simulation results.}

\begin{figure}[htp]
    \centering
 \includegraphics[width=0.45\textwidth]{./Fig_thres0.25.pdf}
  \includegraphics[width=0.45\textwidth]{./Fig_thresh0.9.pdf}
     \caption{\add{Evolution of a randomly initialized network for various powers of $m$ and $n$ in Eq. \eqref{eq:thresh} with a small threshold value of $\beta = 0.25\langle q^m/r^n\rangle $ (left figure), and a large threshold value of $\beta = 0.9\langle q^m/r^n\rangle$ at $t=0$. The network is randomly initialized with $50\times 50$ randomly distributed pores. The black line shows the theoretical transition boundary between channelization instability and homogenization obtained using local dynamics model, i.e., $4m-n=0.75$ (left figure), and $4m-n=0.1$ (right figure).} }\label{fig:figthresh1}% 
\end{figure}

% \begin{figure}[htp]
%     % \centering
%  \includegraphics[width=0.72\textwidth]{Fig_thres0.25.pdf}
%      \caption{\textcolor{blue}{Evolution of a randomly initialized network for various powers of $m$ and $n$ in Eq. \eqref{eq:thresh} with a small thresholding value of $\beta = 0.25\langle q^m/r^n\rangle $ at $t=0$. The network is randomly initialized with $50\times 50$ randomly distributed pores. The black line shows the theoretical transition boundary between channelization instability and homogenization obtained using simplified model, i.e., $m=(n+1)/4$.} }\label{fig:figthresh1}% 
% \end{figure}

% \begin{figure}[htp]
%     % \centering
%  \includegraphics[width=0.72\textwidth]{Fig_thresh0.9.pdf}
%      \caption{\textcolor{blue}{Evolution of a randomly initialized network for various powers of $m$ and $n$ in Eq. \eqref{eq:thresh} with a large thresholding value of $\beta = 0.9\langle q^m/r^n\rangle $ at $t=0$. The network is randomly initialized with $50\times 50$ randomly distributed pores. The black line shows the theoretical transition boundary between channelization instability and homogenization obtained using simplified model, i.e., $m=(n+1)/4$.}}\label{fig:figthresh2}% 
% \end{figure}

@@

% \section{Generalization of the toy model}
% %
% A general erosion in the edges of a network can be written as 
% %
% \begin{align}
%     \frac{dr}{dt} = f(q,r), \label{eq:general}
% \end{align}
% %
% where $q$, and $r$ are the fluid flux and the radius of an edge, and $f(\cdot)$ represents the function that determines the physics of erosion. Note that if $f(q,r)>0$ the edge is eroding, and if $f(q,r)<0$, the edge is clogging. In this section, we study the fate of our toy model, given such general erosion law.

% \textbf{Parallel Tubes: }
% %
% We first consider parallel edges with an arbitrary erosion function, i.e., Eq.~\eqref{eq:general}. In the parallel case, both tubes experiencing the same pressure difference $\delta p$, and using Darcy's law the flux in each tube $q_i = \beta r_i^4 \delta p$, where $\beta = \pi/8\mu L$, $L$ being the length of the tube, $\mu$ the viscosity of the fluid, and $i=1,2$ correspond to the edges. We define $g(r_i) = f(q_i, r_i)$ and as a result the erosion dynamics can be written as $dr_i/dt = g(r_i)$. Without loss of generality we assume $r_1>r_2$. The fluid flux in the tubes can be obtained using $q_i/q = \tilde C_i $, where $q$ is the total flux, and $\tidle C_i = C_i/(C_1+C_2)$ is the relative conductivity, and $C_i = \pi r_i^4/8\mu L $ is the tubes conductance. In order to find the evolution of the fluid flux, we find the time derivative of relative conductivity,  
% %
% \begin{align}
%     \frac{d\tilde{C}_1}{dt} = \frac{4C_1^{\frac{3}{4}} C_2^{\frac{3}{4}}}{(C_1+C_2)^2}(r_2 g(r_1)-r_1 g(r_2)).
% \end{align}
% %
% We find that $r_1=r_2$ is a fixed point, where flow is equally distributed between the edges resulting in homogenization. If $d \tilde C_1/dt <0$, then the edge with larger radius has a reduced growth the smaller radius edge has an increasing growth until $r_1=r_2$ is reached. This condition results in the homogenization where both edge's radius increases until the fixed point of $r_1=r_2$ is reached. The condition of $d \tilde C_1/dt<0$ is satisfied when 
% %
% \begin{align}
%     \frac{g(r_1)}{g(r_2)}<\frac{r_1}{r_2}, \forall r_1>r_2>0.
% \end{align}
% %
% The above result can further be simplified as if $d\left(g(r)/r\right)/dr<0$ which ultimately results in homogenization of the parallel edges. Lastly, we can also study the perturbation of tube's radius around the fixed point as $r_1=r_0_\delta r$ and $r_2 = r_0$. Using Taylor series expansion, we can show that stability condition reduces to $g(r_0)>g'(r_0) r_0$. Therefore, if the erosion rate is sub-linear locally near the initial radius, the network moves towards homogenization in short time, while the long term behavior is less predictable. On the other hand, if the erosion rate is sup-linear, we expect the parallel tubes to move away from homogenization (i.e., channelization) in short times. Applying the power law erosion model with $f(q,r)=\alpha q^m/r^n$, it can be shown that $g(r) \propto  r^{4m-n} $, and therefore homogenization condition becomes $4m-n<1$ as seen in simulations. Although the result is derived under the constant pressure difference assumption, it holds validity for homogenization cases, where the pressure difference is going towards homogenization as well. However, for channelization cases, the pressure difference can be varied vastly inside the network, thus once the channelization happens, negative feedback effect of erosion rate might not be able to reverse that.

% \textbf{Series Tubes: }
% %
% In the series toy model, we have $q_1=q_2 = q = \frac{C_1 C_2}{C_1+C_2}\delta p$ and similar results can be derived so that the condition for homogenization becomes
% %
% \begin{align}
%     \frac{f(q,r_1)}{f(q,r_2)}<\frac{r_1}{r_2}, \forall r_1>r_2>0.
% \end{align}
% %
% Note that if the dependence of $f$ on $q$ is non-homogeneous, then this condition might be valid for some $q$'s but not the others. Applying the power law erosion model with $f(q,r)=\alpha q^m/r^n$, it can be shown that homogenization condition is $n>0$ as it has been seen in simulations.

\add{\section{General shear dependent erosion model}}
\add{
As an example of a nonlinear erosion model, we consider a shear-dependent erosion
    \frac{dr}{dt} = f(\tau),
where $\tau=\tau(q,r) = 4\mu q/\pi r^3$ is the shear stress at the tube's wall assuming Poiseuille flow. It is to be noted that a shear-dependent erosion model is commonly used in studying erosion  \cite{jager2017channelization,bonelli2011micromechanical,parker2000purely} wherein such models erosion happens linearly proportional to the shear stress at the wall only after a certain threshold value $\tau_0$ (see main text for more information). Such linear models, however, ignore the maximum detachment rate of particles at the pore's boundary which would effectively limit the pore's radius maximum rate of change. To address this, we use a nonlinear erosion model
    f(\tau) = \alpha \left( \tanh{\left[ \beta(\tau - \tau_0)\right]}+1\right),
    \label{eq:sigmoidErosion}
where $\beta$ determines the shear rate scale and $\alpha$ determines the erosion rate scale. Note that the erosion rate scale $\alpha$ only affects the timescale at which the final state is reached and has no effect on the pattern formation. Without loss of generality we set $\alpha = 1$. Following the local dynamics model discussion in the main text, one can find that the homogenization condition reduces to 
    f(\tau)>\tau \frac{df(\tau)}{d \tau}.
    \label{eq:sigmoidCondition}
Assuming a random network (Fig.~\ref{SIfig:fig9-sigmoidDelaunay}a), the initial values of shear stress at the walls have a  decaying distribution over a large set of values (Fig.~\ref{SIfig:fig9-sigmoidDelaunay}b). 
%Assuming nonlinear shear dependent erosion model (Eq. \eqref{eq:sigmoidErosion}) three regions of erosion can be identified: (i) The constant erosion region as $\tau\to\infty$. In this region the erosion happens at a constant rate $\lim_{\tau \to \infty} f(\tau) = 2$ and additionally the homogenization condition (Eq.~\eqref{eq:sigmoidCondition}) is satisfied (since $\lim_{\tau \to \infty} f(\tau) = 2$ and $\lim_{\tau \to \infty} f'(\tau) = 0$). It is to be noted that the homogenization effect for the constant erosion rate is similar to $m=0,n=0$ case for the erosion model discussed in the main text; (ii)  The slow-erosion region as $\tau\to 0$. In this region both $f(\tau)$ and $\tau f'(\tau)$ are close to 0 and since $f(\tau)$ is bounded in-between 0 and 2, erosion in the edges at this region happens at a very slower rate compared to other edges with nonzero shear-rate; (iii) The finite erosion-rate region near $\tau = \tau_0$, where $f(\tau)$ is an increasing function. In this region $\tau f'(\tau)$ has a maximum at $\beta \tau_0$ with a bandwidth of $1/\beta$. If $\beta \tau_0 < 1$, then the maximum $\tau f'(\tau)$ does not exceed 1, and the homogenization condition holds true for any $\tau$. As the value of $\tau_0$ increases, the homogenization region moves towards larger values of $\tau$ where edges with smaller values of $\tau$ would result in channelization.  It can be seen that varying $\beta$ and $\tau_0$ changes the network behavior. 
To test our local dynamics model, we vary $\beta$ and $\tau_0$ and run three different tests: (i) $\beta=1,\tau_0=1$: For such values of $\beta,\tau$, almost all the tubes are in the homogenization region corresponding to Eq.~\eqref{eq:sigmoidCondition} (Fig.~\ref{SIfig:fig9-sigmoidDelaunay}c) and the network moves toward homogenization (Fig.~\ref{SIfig:fig9-sigmoidDelaunay}d). The final PDF of the network after erosion is further shown in (Fig.~\ref{SIfig:fig9-sigmoidDelaunay}e). (ii) $\beta=1, \tau_0=12$: In this case, the homogenization condition is satisfied at some finite value of shear stress $\tau_0$ where $f(\tau_0) = \tau_0 f(\tau_0)$. As a result, some edges would homogenize, and some would channelize (Fig.~\ref{SIfig:fig9-sigmoidDelaunay}f). Running the simulation over a random network, we find that a finite number of channels form (Fig.~\ref{SIfig:fig9-sigmoidDelaunay}g). The PDF distribution of shear stress at the wall $\tau$ correspondingly becomes bimodal (Fig.~\ref{SIfig:fig9-sigmoidDelaunay}h). (iii) $\beta=1, \tau_0=30$: In this case, almost all the edges in the network are in the channelization regime  (Fig.~\ref{SIfig:fig9-sigmoidDelaunay}i), and performing the simulations we find that indeed one strong single channel is formed (Fig.~\ref{SIfig:fig9-sigmoidDelaunay}j) and the PDF of shear stress $\tau$ shows a clear separation between flow in the channel versus flow in the rest of the system (Fig.~\ref{SIfig:fig9-sigmoidDelaunay}k). 
}
   \begin{figure}[htp]
     % \centering
      \includegraphics[width = 0.8\textwidth]{sigmoidDelaunay.pdf}
     \caption{\textcolor{blue}{(a) A 2D random network with $N_x=50$ and $N_y=50$ points. (b) The probability distribution function (PDF) of the shear stress for the edges of the network. (c)/(f)/(i) The PDF of the shear stress along with nonlinear erosion law $f(\tau)$ (solid red line) and $\tau f'(\tau)$ (dashed red line). The areas where the homogenization condition (i.e., $f(\tau)>\tau f'(\tau)$) is satisfied are shaded with green color, otherwise, the channelization condition is satisfied and the region is shaded with red color. (d)/(g)/(j) Snapshots of the 2D network after erosion using the nonlinear erosion law shown in the previous row until the average radius of the network is increased to twice its original size. (e)/(h)/(k) The PDF of the shear stress at the walls for the final snapshot of the network shown in the previous row. 
    }}
     \label{SIfig:fig9-sigmoidDelaunay}

%  \begin{figure}[htp]
%      % \centering
%       \includegraphics[width = 1.0\textwidth]{sigmoidDelaunay_rev1.pdf}
%      \caption{\add{Nonlionear erosion in a 2D random network of tubes with $N_x=50$ and $N_y=50$, where edge radii are sampled from a uniform distribution with $\mathcal{U}[1,14]$. Snapshots of the network after for different erosion rate $\beta$ and threshold values $\tau_0$ are shown after $N$ steps such that $\langle r\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle = 7.5$. The erosion law is based on Eq. \eqref{eq:sigmoidErosion}. As shown above, increasing $\tau_0$ results in a stronger channelization, while changing $\beta$ has a much smaller effect. Note that increasing $\tau_0$, increases the number of tubes in the channelization region, while increasing $\beta$ slightly shifts the region and has a much smaller effect.} }
%      \label{SIfig:fig10-sigmoidDelaunay}
%  \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \begin{align}
%     \frac{dr_i}{dt} = f(|q_i|,r_i),
% \end{align}
% %
% where $i = 1$ and $2$ correspond to pipe 1 and pipe 2 respectively and $f(q,r)$ can be negative thus it allows clogging. In the toy model case $q_i = C p r_i^4$ where $p$ is the pressure difference between two terminals of the pipes. Therefore the erosion rate only depends on $r_i$
% \begin{align}
%     f(|q_i|,r_i) = g(r_i).
% \end{align}
% The time derivative of relative conductivity $\tilde{C}_1 = C_1/(C_1+C_2)$ can be written as
% \begin{align}
%     \frac{d\tilde{C}_1}{dt} = \frac{4C_1^{\frac{3}{4}} C_2^{\frac{3}{4}}}{(C_1+C_2)^2}(r_2 g(r_1)-r_1 g(r_2)).
% \end{align}
% It is obvious that in the toy model, $r_1=r_2$ is a fixed point, while its stability depends on the form of $f(q,r)$, or equivalently $g(r)$ in the parallel case. It can be shown that the condition for homogenization is 
% \begin{align}
%     \frac{g(r_1)}{g(r_2)}<\frac{r_1}{r_2}, \forall r_1>r_2>0.
% \end{align}
% This is a strong constraint to $g(r)$, and in reality it might not hold true for arbitrary $r$'s when negative feedback over erosion rate is introduced. One generalization we can make is to expand the function near the value we are interested in. Let $r_0 = r_2$ be the initial value, and $r_1 = r_0 + \delta r$ to be the perturbation in one pipe. With Taylor series expansion the above condition can be written as
% \begin{align}
%     g(r_0)>g'(r_0) r_0.
% \end{align}
% Therefore, if the erosion rate is sub-linear locally near the initial radius, we can say it will drive the network towards homogenization in short time, while the long term behavior is less predictable. On the other hand, if the erosion rate is sup-linear, we expect to see channelization to some extent.

% Applying the power law erosion model with $f(q,r)=C q^m/r^n$, it can be shown that $g(r) = C p r^(4m-n) $ and therefore homogenization condition is $4m-n<1$ as seen in simulations.

% Although the result is derived under the constant pressure difference assumption, it holds validity for homogenization cases, where the pressure difference is going towards homogenization as well. However, for channelization cases, the pressure difference can be varied vastly inside the network, thus once the channelization happens, negative feedback effect of erosion rate might not be able to reverse that.

% In the series toy model, we have $q_1=q_2 = q = p \frac{C_1 C_2}{C_1+C_2}$ and similar results can be derived so that the condition for homogenization is
% \begin{align}
%     \frac{f(q,r_1)}{f(q,r_2)}<\frac{r_1}{r_2}, \forall r_1>r_2>0.
% \end{align}
% Note that if the dependence of $f$ on $q$ is non-homogeneous, then this condition might be valid for some $q$'s but not the others.

% Applying the power law erosion model with $f(q,r)=C q^m/r^n$, it can be shown that homogenization condition is $n>0$ as seen in simulations.

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \bibliography{ref}% Produces the bibliography via BibTeX.
%apsrev4-2.bst 2019-01-14 (MD) hand-edited version of apsrev4-1.bst
%Control: key (0)
%Control: author (8) initials jnrlst
%Control: editor formatted (1) identically to author
%Control: production of article title (0) allowed
%Control: page (0) single
